### 2018-12-09 Revision/Response/response.tex

\item \referee{6.38: typo, $w_r$ should be $w_1$}

\noindent  The typo has been fixed in the revised manuscript.

\item \referee{7.23-27: why 'statistical independence' and not just 'independence'? Is there any other 
type of independence?}

\item \referee{7.29-30: Any PDF can be written in a Boltzmann form, V(x) is just the negative log-PDF, 
I am not sure if it is necessary to bring up a Boltzmann distribution language here.
Also, it could be confused with Maxwell-Boltzmann distribution, which is a very specific PDF form.}
Thanks for this comment. We followed the theory in ref [27] in the manuscript, 
where the authors derive Poincare type inequalities for functions of 
iid random inputs distributed according to Boltzmann probability laws. We point
out that the uniform distribution does not belong to this class. That was the reason 
for assuming the distribution laws of the
random inputs are either \emph{uniform} or belong to 
the Boltzmann class, in section 3. An important class of Boltzmann measures are 
the so called log-concave distributions, such as normal, exponential, and beta, and gamma
distributions.

We share the reviewer's concern about the potential confusions some readers might face when 
seeing the mention of Boltzmann. We have removed the mention of Boltzmann measures
from the introduction, and now only refer to this in Section 3, where the theoretical
results are discussed.

### 2018-12-09 Revision/Paper/app.tex
..

Secondly, we verified the accuracy of the two surrogates in a probabilistic setting. In particular, we compared 
probability density functions (PDFs) obtained using the true set of model evaluations, and 1-dimensional
surrogates ($\hat{G}$'s) based on the two approaches for gradient estimation as shown in 
Figure~\ref{fig:pdf_36D}. Note that the three PDFs were evaluated using the same set of 10$^4$ samples in the 
cross-validation set. 
%
\begin{figure}[htbp]
\begin{center}
\includegraphics[width=3.0in]{./Figures/pdf_plot36Dp2}
\end{center} 
\caption{\rebut{A comparison of the PDFs of ignition delay, obtained using model 
evaluations (solid line) and 1-dimensional surrogates using the regression-based strategy (dashed line) and the
perturbation-based strategy (dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were 
used in each case.}}
\label{fig:pdf_36D}
\end{figure}
%
\rebut{The PDFs are observed to be in close agreement with each other. Specifically, the modal
estimate and the uncertainty (quantified by the spread in the distributions) is found to be consistent for the three 
cases. To confirm this, we compute the first-order (mean) and the second-order (standard deviation) statistics
of the estimates of the ignition delay obtained using the model, 1-dimensional surrogate from perturbation, and
1-dimensional surrogate from regression at the cross-validation sample set. The mean and standard deviation
estimates are provided in Table~\ref{tab:stats}.}
%
\begin{table}[htbp]
\begin{center}
\begin{tabular}{ccc}
\toprule
$\textbf{Distribution}$ & $\mu$ & $\sigma$ \\ 
\bottomrule
$G$~(Model) & \rebut{0.1510} & \rebut{0.1373} \\
$\hat{G}$~(Perturbation-based) & \rebut{0.1514} & \rebut{0.1275} \\
$\hat{G}$~(Regression-based) & \rebut{0.1510} & \rebut{0.1330} \\
\bottomrule
\end{tabular}
\caption{The mean ($\mu$), and the standard deviation ($\sigma$), computed using the model ($G$), and
the surrogate ($\hat{G}$) based on the two strategies at 10$^4$ samples in the cross-validation
set.}
\label{tab:stats}
\end{center}
\end{table}
%
\rebut{The mean and the standard deviation estimates obtained using the model and the 1-dimensional surrogates
are found to be in close agreement. It is interesting to note that estimates from the regression approach are
observed to be slightly more accurate. This is due to the fact that the converged active subspace in this case is 
based on 1850 samples in the 36-dimensional input domain, whereas, the converged active subspace based the 
perturbation approach is based on only 25 samples. As a result, the surrogate based on regression explores
a much wider range of values of the ignition delay compared to the surrogate based on perturbation as shown in
Figure~\ref{fig:hd}~(right).}

>>>>

Secondly, we verified the accuracy of the two surrogates in a probabilistic setting. In particular, we compared 
probability density functions (PDFs) obtained using the true set of model evaluations, and 1-dimensional
surrogates ($\hat{G}$'s) based on the two approaches for gradient estimation as shown in 
Figure~\ref{fig:pdf_36D}. Note that the three PDFs were evaluated using the same set of 10$^4$ samples in the 
cross-validation set. 
%
\begin{figure}[htbp]
\begin{center}
\includegraphics[width=3.0in]{./Figures/pdf_plot36Dp2}
\end{center} 
\caption{\rebut{A comparison of the PDFs of ignition delay, obtained using model 
evaluations (solid line) and 1-dimensional surrogates using the regression-based strategy (dashed line) and the
perturbation-based strategy (dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were 
used in each case.}}
\label{fig:pdf_36D}
\end{figure}
%
\rebut{The PDFs are observed to be in close agreement with each other. Specifically, the modal
estimate and the uncertainty (quantified by the spread in the distributions) is found to be consistent for the three 
cases. To confirm this, we compute the first-order (mean) and the second-order (standard deviation) statistics
of the estimates of the ignition delay obtained using the model, 1-dimensional surrogate from perturbation, and
1-dimensional surrogate from regression at the cross-validation sample set. The mean and standard deviation
estimates are provided in Table~\ref{tab:stats}.}
%
\begin{table}[htbp]
\begin{center}
\begin{tabular}{ccc}
\toprule
$\textbf{Distribution}$ & $\mu$ & $\sigma$ \\ 
\bottomrule
$G$~(Model) & \rebut{0.1510} & \rebut{0.1373} \\
$\hat{G}$~(Perturbation-based) & \rebut{0.1514} & \rebut{0.1275} \\
$\hat{G}$~(Regression-based) & \rebut{0.1510} & \rebut{0.1330} \\
\bottomrule
\end{tabular}
\caption{The mean ($\mu$), and the standard deviation ($\sigma$), computed using the model ($G$), and
the surrogate ($\hat{G}$) based on the two strategies at 10$^4$ samples in the cross-validation
set.}
\label{tab:stats}
\end{center}
\end{table}
%
\rebut{The mean and the standard deviation estimates obtained using the model and the 1-dimensional surrogates
are found to be in close agreement. Hence, although the relative L$^2$ norm of the discrepancy~($\varepsilon_d$)
is found to be considerably large, the uncertainty in the ignition delay is captured by both 
approaches with reasonable accuracy. 
It is however interesting to note that estimates from the regression approach are
observed to be slightly more accurate. This is potentially due to the fact that the 
converged active subspace in this case is 
based on 1850 samples in the 36-dimensional input domain, whereas, the converged active subspace based the 
perturbation approach is based on only 25 samples. As a result, the surrogate based on regression explores
a much wider range of values of the ignition delay compared to the surrogate based on perturbation as shown in
Figure~\ref{fig:hd}~(right).}

### 2018-12-09 Revision/Paper/method.tex
..

The accuracy of the two approaches was assessed by estimating $\varepsilon_{\text{L}^{2}}^{N-n}$ 
using the components of the dominant eigenvector in the converged active subspace in each case
in~\eqref{eq:accu}. The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ was estimated to be 0.0657
and 0.1050 using perturbation and regression respectively. Hence, the regression approach was
found to be relatively more accurate. Squared values of the individual components of the 
dominant eigenvector from the
two approaches and for the case using $N$~=~1000 in the perturbation approach are plotted in
Figure~\ref{fig:comp}.}
%
%is plotted in
%Figure~\ref{fig:comp} (center) to illustrate convergence characteristics based on Algorithm~\ref{alg:grad} 
%for the two approaches. As observed for a set tolerance $\tau$~=~0.05, 8 iterations are needed 
%to converge in both cases. 
%At each iteration, 5 new random samples were added for the perturbation approach. As discussed earlier, gradient 
%computation (using finite difference) at 5 samples requires 100 model runs. Hence, in order to compare the
%two approaches for the same computational effort, we generate 100 random samples in the 19D input domain for 
%the regression approach. Therefore, 8 iterations require 800 model runs leading to a total of 900 model runs 
%including the initial step in both cases.  The accuracy of the two approaches is compared for the same 
%computational effort by computing $\varepsilon_{\text{L}^{2}}^{N-n}$ using~\eqref{eq:accu} in 
%Figure~\ref{fig:comp} (right). Once again, $N$
%corresponds to the case of active subspace computation using the perturbation approach using 1000 samples
%in the 19D domain; $n$ ranges from 5 to 45 for the perturbation approach, and 100 to 900 for the regression 
%approach corresponding to iterations 1 to 8. The accuracy of the two approaches is found to be comparable for
%the same amount of computational effort. However, the perturbation approach is observed to be slightly more
%accurate.}
%
%\begin{figure}[htbp]
% \begin{center}
%  \includegraphics[width=0.8\textwidth]{./Figures/eig_convD19}
%\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
%eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
%Additionally, the dominant eigenvector components obtained using $M$~=~20000 model runs (corresponding to 
%$N$~=~1000 samples)
%in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.
%Center: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
%is plotted for successive iterations to illustrate convergence behavior for the two strategies. Right: The
%quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ is plotted as a function of iterations to compare the accuracy
%of the two strategies. Note that the above plots compare the convergence and accuracy of the two
%strategies for the same amount of computational effort, quantified by the number of model runs, $M$.}}
%\label{fig:comp}
%\end{center}
%\end{figure}
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_conv19D}
\caption{\rebut{An illustrative comparison of individual squared components of the converged dominant
eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
Additionally, the dominant eigenvector components obtained using $M$~=~20000 model runs (corresponding to 
$N$~=~1000 samples)
in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.}}
\label{fig:comp}
\end{center}
\end{figure}
%
\rebut{
The set of eigenvector components for the three cases are found to be in excellent agreement with
each other, indicating that both
approaches have sufficiently converged and are reasonably accurate for this setup.}

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the ignition delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in \rebut{a 1-dimensional} active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the \rebut{1-dimensional active} subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
\caption{\rebut{An illustrative comparison of the SSPs generated using the 
perturbation and the regression strategies for computing the active subspace.}}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
\rebut{The dominant eigenvector obtained using perturbation is based on $N$~=~45 samples which requires
$M$~=~900 model runs. For the same amount of computational effort, we can afford $N$~=~900 samples
when using regression. Hence, the SSP from regression is based on 900
points: ($\hat{\mat{W}}_1^\top\bm{\xi}$,~$G(\hat{\mat{W}}_1^\top\bm{\xi})$). On the other hand, the SSP
from perturbation is plotted using only 45 points as mentioned earlier.
Nevertheless, the illustrative comparison clearly indicates that the two SSPs
are in excellent agreement. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $\log(A_i)$ although non-linear, can be approximated by
a 1-dimensional active subspace.}

>>>>

The accuracy of the two approaches was assessed by estimating $\varepsilon_{\text{L}^{2}}^{N-n}$ 
using the components of the dominant eigenvector in the converged active subspace in each case
in~\eqref{eq:accu}. The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ was estimated to be 0.0657
and 0.1050 using perturbation and regression respectively. Hence, the regression approach was
found to be relatively more accurate. Squared values of the individual components of the 
dominant eigenvector from the
two approaches and for the case using $N$~=~1000 in the perturbation approach are plotted in
Figure~\ref{fig:comp}~(left).}
%is plotted in
%Figure~\ref{fig:comp} (center) to illustrate convergence characteristics based on Algorithm~\ref{alg:grad} 
%for the two approaches. As observed for a set tolerance $\tau$~=~0.05, 8 iterations are needed 
%to converge in both cases. 
%At each iteration, 5 new random samples were added for the perturbation approach. As discussed earlier, gradient 
%computation (using finite difference) at 5 samples requires 100 model runs. Hence, in order to compare the
%two approaches for the same computational effort, we generate 100 random samples in the 19D input domain for 
%the regression approach. Therefore, 8 iterations require 800 model runs leading to a total of 900 model runs 
%including the initial step in both cases.  The accuracy of the two approaches is compared for the same 
%computational effort by computing $\varepsilon_{\text{L}^{2}}^{N-n}$ using~\eqref{eq:accu} in 
%Figure~\ref{fig:comp} (right). Once again, $N$
%corresponds to the case of active subspace computation using the perturbation approach using 1000 samples
%in the 19D domain; $n$ ranges from 5 to 45 for the perturbation approach, and 100 to 900 for the regression 
%approach corresponding to iterations 1 to 8. The accuracy of the two approaches is found to be comparable for
%the same amount of computational effort. However, the perturbation approach is observed to be slightly more
%accurate.}
%\begin{figure}[htbp]
% \begin{center}
%  \includegraphics[width=0.8\textwidth]{./Figures/eig_convD19}
%\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
%eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
%Additionally, the dominant eigenvector components obtained using $M$~=~20000 model runs (corresponding to 
%$N$~=~1000 samples)
%in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.
%Center: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
%is plotted for successive iterations to illustrate convergence behavior for the two strategies. Right: The
%quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ is plotted as a function of iterations to compare the accuracy
%of the two strategies. Note that the above plots compare the convergence and accuracy of the two
%strategies for the same amount of computational effort, quantified by the number of model runs, $M$.}}
%\label{fig:comp}
%\end{center}
%\end{figure}
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.38\textwidth]{./Figures/eig_conv19D}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
Additionally, the dominant eigenvector components obtained using $M$~=~20000 model runs (corresponding to 
$N$~=~1000 samples)
in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.
Right: An illustrative comparison of the SSPs generated using the
perturbation and the regression strategies for computing the active subspace. 
}}
\label{fig:comp}
\end{center}
\end{figure}
\rebut{
The set of eigenvector components for the three cases are found to be in excellent agreement with
each other, indicating that both
approaches have sufficiently converged and are reasonably accurate for this setup.}

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the ignition delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in \rebut{a 1-dimensional} active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the \rebut{1-dimensional active} subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp}~(right).
%
%\begin{figure}[htbp]
% \begin{center}
%  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
%\caption{\rebut{An illustrative comparison of the SSPs generated using the 
%perturbation and the regression strategies for computing the active subspace.}}
%\label{fig:comp_ssp}
%\end{center}
%\end{figure}
%
\rebut{The dominant eigenvector obtained using perturbation is based on $N$~=~45 samples which requires
$M$~=~900 model runs. For the same amount of computational effort, we can afford $N$~=~900 samples
when using regression. Hence, the SSP from regression is based on 900
points: ($\hat{\mat{W}}_1^\top\bm{\xi}$,~$G(\hat{\mat{W}}_1^\top\bm{\xi})$). On the other hand, the SSP
from perturbation is plotted using only 45 points as mentioned earlier.
Nevertheless, the illustrative comparison clearly indicates that the two SSPs
are in excellent agreement. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $\log(A_i)$ although non-linear, can be approximated by
a 1-dimensional active subspace.}
